Modal Abstractions for Virtualizing Memory Addresses

### 2023-11-16 programlogic.tex
clean up framing discussion

This is then how we justify our actual rule for \lstinline|cr3| updates:
\todo[inline]{Ismail: Can we chat Wednesday about this rule, which has cr3 register pointstos instead of the cr3 subscripts?
The subscript form might make the derivation of a frame rule harder.}
\begin{mathpar}
\inferrule*[left=ChangeAddressSpace]{
  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\overline{is}
}{
  \{P \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast [r_2](Q)\}
  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
  %\lstinline|mov %cr3, r| 
}
\end{mathpar}
Because the precondition on this rule is global, we avoid issues with framing.

If we wanted to consider a frame rule that would work for this logic, we could consider:
\begin{mathpar}
  \inferrule*[left=Cr3Frame]{
    \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
  }{
    \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
  }
\end{mathpar}
By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally. In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
Attempting to derive a frame rule as some work using Hoare doubles does~\cite{Chlipala2013Bedrock}
would enforce this restriction.

Another useful perspective on this might be the dynamic frames~\cite{parkinson2011relationship} 
view: virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
% Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.

>>>>

This is then how we justify our actual rule for \lstinline|cr3| updates, which is \emph{almost}
\begin{mathpar}
\inferrule*[left=ChangeAddressSpace]{
  \{[r_1](P) \ast cr3 \mapsto_{r} r_2 \ast r \mapsto_{r} r_2 \ast Q\}\;\overline{is}
}{
  \{P \ast cr3 \mapsto_{r} r_1 \ast r \mapsto_{r} r_2 \ast [r_2](Q)\}\;
  \texttt{mov}~\texttt{\%cr3},~r;\;\overline{is}
  %\lstinline|mov %cr3, r| 
}
\end{mathpar}
Because the precondition on this rule is global, we avoid issues with framing. A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock} A more typical frame rule can be recovered from Hoare doubles in the typical way~\cite{Chlipala2013Bedrock}
with the added restriction that \lstinline|cr3| is restored to its pre-framing value before un-framing (in this case, the framed-out assertion is effectively
stashed under an other-space assertion for the initial address space, and pulled out as part of restoring).

% If we wanted to consider a frame rule that would work for this logic, we could consider:
% \begin{mathpar}
%   \inferrule*[left=Cr3Frame]{
%     \{P\ast cr3=v\}\;C\;\{ Q \ast cr3=v\}
%   }{
%     \{R\ast P\ast cr3=v\}\;C\;\{ R \ast Q \ast cr3=v\}
%   }
% \end{mathpar}
% By demanding that \lstinline|cr3| be held constant (or rather, at least restored to its original value) we could frame almost traditionally. In particular, this rule would work with framing around calls that might lead to address space switches, such as calling blocking operations in the kernel.
% Attempting to derive a frame rule as some work using Hoare doubles does~\cite{Chlipala2013Bedrock}
% would enforce this restriction.

% Another useful perspective on this might be the dynamic frames~\cite{parkinson2011relationship} 
% view: virtual points-to assertions are self-stable \emph{except} for changes in \lstinline|cr3|, so framing would then naturally require other means of holding \lstinline|cr3| constant (or saving and restoring it).
% Virtual points-to assertions could be made self-stable by also giving them partial ownership over \lstinline|cr3| assertions, but this would require explicitly plumbing that ownership from \emph{all} assertions back to any place an address space change might occur; this would seem to be a far graver loss of modularity than this extra quirk in framing discussions.
